Distributed Algorithms for Sparse Matrix-Vector Multiplication

### 2015-05-01 paper.Rmd
More paper updates

1. Load-Balanced Sparse Matrix–Vector Multiplication on Parallel Computers by Nastea, Frieder, and El-Ghazawi
2. An architecture-aware technique for optimizing sparse matrix-vector multiplication on GPUs by Maggioni and Berger-Wolf
3. A model-driven blocking strategy for load balanced sparse matrix–vector multiplication on GPUs

@@

```{r, warning=FALSE}
wearables <- read.csv("data/Example_WearableComputing_weight_lifting_exercises_biceps_curl_variations.csv")
X <- apply(wearables[,-1], 2, as.numeric)
y <- as.numeric(wearables[,1])

@@

\caption{Overview of expected coverage for $k$-peptide libraries of different sizes $N$ with the different encoding schemes (NNN, NNB, NNK/S, and trimer). An additional line for maximum possible coverage is shown.}
\label{fig:alg1}
\end{figure}

### Results
The performance improvements were evaluated using four different matrices of slightly different characteristics, and performing the simulation on an Intel Paragon supercomputer. Figure \ref{fig:im1} gives a table describing the different properties of the test matrices. Note the sparsity value is given as the proportion of non-zero elements. Hence, all of the matrices test are quite sparse, containing at most 5.55 non-zero elements.

@@

\caption{Overview of expected coverage for $k$-peptide libraries of different sizes $N$ with the different encoding schemes (NNN, NNB, NNK/S, and trimer). An additional line for maximum possible coverage is shown.}
\label{fig:im1}
\end{figure}

@@

\caption{Overview of expected coverage for $k$-peptide libraries of different sizes $N$ with the different encoding schemes (NNN, NNB, NNK/S, and trimer). An additional line for maximum possible coverage is shown.}
\label{fig:im2}
\end{figure}

@@

\caption{Overview of expected coverage for $k$-peptide libraries of different sizes $N$ with the different encoding schemes (NNN, NNB, NNK/S, and trimer). An additional line for maximum possible coverage is shown.}
\label{fig:im3}
\end{figure}

@@

The object is to minimize $z(S) = \sum_{j=0}^{k-1}|\cap_{i=0}^{w-1} S_{i,j}|$. In other words, the goal is to minimize the sum of cache line intersections in the scheduling table S (as these are redundant transactions). In Figure \ref{fig:im5}, the results of the algorithm are given in a table. The authors note that the speedup is not significant, but that novel contributions in this arena have not performed significantly better.

@@

\caption{Overview of expected coverage for $k$-peptide libraries of different sizes $N$ with the different encoding schemes (NNN, NNB, NNK/S, and trimer). An additional line for maximum possible coverage is shown.}
\label{fig:im5}
\end{figure}

@@

\caption{Overview of expected coverage for $k$-peptide libraries of different sizes $N$ with the different encoding schemes (NNN, NNB, NNK/S, and trimer). An additional line for maximum possible coverage is shown.}
\label{fig:dmpc}
\end{figure}

@@

\caption{Overview of expected coverage for $k$-peptide libraries of different sizes $N$ with the different encoding schemes (NNN, NNB, NNK/S, and trimer). An additional line for maximum possible coverage is shown.}
\label{fig:bil}
\end{figure}

>>>>

1. Load-Balanced Sparse Matrix–Vector Multiplication on Parallel Computers by Nastea, Frieder, and El-Ghazawi
2. An architecture-aware technique for optimizing sparse matrix-vector multiplication on GPUs by Maggioni and Berger-Wolf
3. Scalable Parallel Matrix Multiplication on Distributed Memory Parallel Computers by Li

@@

```{r, warning=FALSE}
wearables <- read.csv("data/Example_WearableComputing.csv")
X <- apply(wearables[,-1], 2, as.numeric)
y <- as.numeric(wearables[,1])

@@

\caption{Pseudo-code for the greedy algorithm of the first paper. The algorithm performs operations on a major node (a leader). The leader uses the GALA routine to assign the most dense remaining rows to the processors with the minimum current workload.}
\label{fig:alg1}
\end{figure}

### Results
The performance improvements were evaluated using four different matrices of slightly different characteristics, and performing the simulation on an Intel Paragon supercomputer. Figure \ref{fig:im1} gives a table describing the different properties of the test matrices. Note the sparsity value is given as the proportion of non-zero elements. Hence, all of the matrices test are quite sparse, containing at most 5.55% non-zero elements.

@@

\caption{Matrices used for the evaluation of the algorithm. In particular, note that the last matrix is a synthetically generated matrix from a Zipf distribution. This distribution simulates sparsity and skewness characteristics that GALA is most effective at handling.}
\label{fig:im1}
\end{figure}

@@

\caption{Graph of the speedup of the algorithm as a function of the number of processors. Note that GALA performs most strongly on the bottom right panel, which is the panel corresponding to the synthetically generated data.}
\label{fig:im2}
\end{figure}

@@

\caption{Sliced ELL compression format for sparse matrices presented in this paper. In particular, since there is a local value for the number of columns (k) for each warp, the sliced ELL format allows smaller matrices to be divided up to each warp locally.}
\label{fig:im3}
\end{figure}

@@

The object is to minimize $z(S) = \sum_{j=0}^{k-1}|\cap_{i=0}^{w-1} S_{i,j}|$. In other words, the goal is to minimize the sum of cache line intersections in the scheduling table S (as these are redundant transactions). The algorithm presented does so by greedily choosing a row which maximizes the amount of overlap with rows that have already been placed, which in turn will minimize the number of cache transactions. The time complexity amoritized is approximately $O(n)$ for this algorithm (technically, it is $O(k^2n)$, but because the sliced ELL format allows k to be significantly smaller in sparse matrices than n, this is approximated well by $O(n)$). In Figure \ref{fig:im5}, the results of the algorithm are given in a table. The authors note that the speedup is not significant, but that novel contributions in this arena have not performed significantly better.

@@

\caption{Table listing the reduction in the number of cache transactions for several different benchmarks. Overall, a 1.09x speedup is noted by using this algorithm.}
\label{fig:im5}
\end{figure}

@@

\caption{A diagram of Distributed Memory Parallel Computers (DMPC). Note that this structure allows for one-to-one processor communication.}
\label{fig:dmpc}
\end{figure}

@@

\caption{The recursive billinear algorithm for matrix multiplication. The algorithm presented in the paper unrolls this algorithm into a parallel iterative algorithm.}
\label{fig:bil}
\end{figure}
